Portable Multi- and Many-Core Performance for Finite Difference Codes; Application to the Free-Surface Component of NEMO.

### 2016-08-31 nemolite2d_perf.tex
nemolite2d-perf: add para on habakkuk and OpenMP to conclusions

In a previous paper~\cite{shallow_psykal} we introduced a possible
approach to tackling this problem which we term \psykal\ (discussed
below). In that work we considered the implications for serial
performance of the extensive code re-structuring required by the
approach when applied to the `Shallow' shallow-water model. We found
that although the re-structuring did initially incur a sizeable
performance penalty, it was possible to transform the resulting code
to recover performance while obeying the \psykal\ separation of
concerns. In this work we move to looking at parallel performance
within the \psykal\ approach.

@@

The advantage of the \psykal\ re-structuring becomes apparent if we
wish to run NEMOLite2D on different hardware, \textit{e.g.} a
GPU. This is because the necessary code modifications are, by design,
limited to the middle PSy layer. In order to port the Fortran
NEMOLite2D to GPU we used OpenACC directives as opposed to the OpenMP
directives used for the multi-core CPU.

@@

In Figure~\ref{FIG_roofline} we show the performance of kernels from
both Shallow and NEMOLite2D on the roofline model XSfor the E5-1620 CPU.
This demonstrates that the Shallow kernel is achieving a performance
roughly consistent with saturating the available memory bandwidth. In
contrast, the kernel taken from NEMOLite2D is struggling to reach a
performance consistent with saturating the bandwidth to main
memory. We experimented with reducing the problem size (so as to
ensure it fitted within cache) but that did not significantly improve
the performance of the kernel. This then points to more fundamental issues
with the way that the kernel is implemented which are not captured in the
simple roofline model.

@@

>>>>

In a previous paper~\cite{shallow_psykal} we introduced a possible
approach to tackling this problem which we term \psykal\ (discussed
below). In that work we considered the implications for serial
performance of the extensive code re-structuring required by the
approach when applied to the `Shallow' shallow-water model. We found
that although the re-structuring did initially incur a sizeable
performance penalty, it was possible to transform the resulting code
to recover performance (for a variety of CPU/compiler combinations)
while obeying the \psykal\ separation of concerns. In this work we
move to looking at portable parallel performance within the
\psykal\ approach.

@@

The advantage of the \psykal\ re-structuring becomes apparent if we
wish to run NEMOLite2D on different hardware, \textit{e.g.} a
GPU. This is because the necessary code modifications are, by design,
limited to the middle PSy layer. In order to port the Fortran
NEMOLite2D to GPU we used OpenACC directives as opposed to the OpenMP
directives used for the multi-core CPU. We also experimented with
using CUDA directly in order to quantify any performance penalty
incurred.

@@

In Figure~\ref{FIG_roofline} we show the performance of kernels from
both Shallow and NEMOLite2D on the roofline model for the E5-1620 CPU.
This demonstrates that the Shallow kernel is achieving a performance
roughly consistent with saturating the available memory bandwidth. In
contrast, the kernel taken from NEMOLite2D is struggling to reach a
performance consistent with saturating the bandwidth to main
memory. We experimented with reducing the problem size (so as to
ensure it fitted within cache) but that did not significantly improve
the performance of the kernel. This then points to more fundamental issues
with the way that the kernel is implemented which are not captured in the
simple roofline model.

@@

Investigation of the absolute serial performance of the NEMLite2D code
using the roofline model revealed that it was still significantly
below any of the traditional roofline ceilings. We have developed a
code-analysis tool~\cite{habakkuk} that is capable of providing more
accurate ceilings by analysing the nature of the floating point
computations performed by a kernel. In future work we aim to extend
this tool such that it can be applied to code parallelised using
OpenMP.
